---
title: SHIP Report
params:
  consoleWidth: 80
output:
  html_document:
    df_print: paged
    theme: journal
    toc: yes
    toc_depth: 4
    toc_float: no
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    toc: yes
    toc_depth: 4

---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=5,
                      fig.path="Figs/",fig.fullwidth=TRUE,
                      warning=FALSE, message=FALSE, comment = "")
options(width = params$consoleWidth)
```

### INTRODUCTION 
This project is a bare bone exploration of the data ship_study.csv. Make the changes as required.

```{r libraries_load}
# Loading the libraries
library(analyzer)
library(dplyr)
library(corrplot)
library(ggplot2)
library(MASS)
```

***

### DATA 
First, let's load the data and take a look at its dimension and first few rows.

```{r data_load}
tb <- read.csv('data/ship_study.csv')

# Defining factor variables 
factor_vars <- c('sex', 'obs_bp', 'obs_soma', 'obs_int', 'school', 'family', 'smoking', 'myocard', 'stroke', 'diab_known', 'contraception', 'income')

# Converting categorical variables into factors
for (i in factor_vars) {
  tb[,i] <- as.factor(tb[,i])
}
```

The dimension of data is
```{r data_dim, echo=F}
cat(paste0('Columns: ', prettyNum(ncol(tb), big.mark = ','), '
Rows: ', prettyNum(nrow(tb), big.mark = ','), '
Unique Rows: ', prettyNum(nrow(unique(tb)), big.mark = ',')))
```

```{r data_head}
head(tb)
```

***


#### MISSING VALUES 
By plotting the proportion of missing values we can see see which variables has the maximum
counts of missing values. From the plot we can see that column 'mpg' has the highest number
of missing values. Rest columns doesn't have any missing values.

```{r NA_plots}
# Plotting the missing values
analyzer::plotNA(tb, row.level = TRUE)
```

In the right plot we can see how missing values (shown in red color) are spread across all the
columns and all the rows.

***


####  VARIABLE EXPLORATION 
In this section all the individual variables are being explored. **No**
variable is selected as the response (or dependent) variable. While
the remaining variables are selected as the explanatory (or independent) variables.

First, let's create and save all the plots:
```{r save_plots_vars}
variable_plots <- plottr(tb, yvar = NULL,
                         yclass = NULL)
```

##### **Variable: id**
```{r variable_id, message = TRUE}
explainer(tb$id)
```

```{r plot_id}
plot(variable_plots$id)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = id)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$id, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: exdate**
```{r variable_exdate, message = TRUE}
explainer(tb$exdate)
```

```{r plot_exdate}
plot(variable_plots$exdate)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = exdate)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$exdate, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: age**
```{r variable_age, message = TRUE}
explainer(tb$age)
```

```{r plot_age}
plot(variable_plots$age)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = age)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$age, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: sex**
```{r variable_sex, message = TRUE}
explainer(tb$sex)
```

```{r plot_sex}
plot(variable_plots$sex)
```

##### **Variable: obs_bp**
```{r variable_obs_bp, message = TRUE}
explainer(tb$obs_bp)
```

```{r plot_obs_bp}
plot(variable_plots$obs_bp)
```

##### **Variable: obs_soma**
```{r variable_obs_soma, message = TRUE}
explainer(tb$obs_soma)
```

```{r plot_obs_soma}
plot(variable_plots$obs_soma)
```

##### **Variable: obs_int**
```{r variable_obs_int, message = TRUE}
explainer(tb$obs_int)
```

```{r plot_obs_int}
plot(variable_plots$obs_int)
```

##### **Variable: dev_bp**
```{r variable_dev_bp, message = TRUE}
explainer(tb$dev_bp)
```

```{r plot_dev_bp}
plot(variable_plots$dev_bp)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = dev_bp)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$dev_bp, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: dev_length**
```{r variable_dev_length, message = TRUE}
explainer(tb$dev_length)
```

```{r plot_dev_length}
plot(variable_plots$dev_length)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = dev_length)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$dev_length, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: dev_weight**
```{r variable_dev_weight, message = TRUE}
explainer(tb$dev_weight)
```

```{r plot_dev_weight}
plot(variable_plots$dev_weight)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = dev_weight)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$dev_weight, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: sbp1**
```{r variable_sbp1, message = TRUE}
explainer(tb$sbp1)
```

```{r plot_sbp1}
plot(variable_plots$sbp1)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = sbp1)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$sbp1, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: sbp2**
```{r variable_sbp2, message = TRUE}
explainer(tb$sbp2)
```

```{r plot_sbp2}
plot(variable_plots$sbp2)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = sbp2)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$sbp2, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: dbp1**
```{r variable_dbp1, message = TRUE}
explainer(tb$dbp1)
```

```{r plot_dbp1}
plot(variable_plots$dbp1)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = dbp1)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$dbp1, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: dbp2**
```{r variable_dbp2, message = TRUE}
explainer(tb$dbp2)
```

```{r plot_dbp2}
plot(variable_plots$dbp2)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = dbp2)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$dbp2, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: height**
```{r variable_height, message = TRUE}
explainer(tb$height)
```

```{r plot_height}
plot(variable_plots$height)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = height)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$height, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: weight**
```{r variable_weight, message = TRUE}
explainer(tb$weight)
```

```{r plot_weight}
plot(variable_plots$weight)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = weight)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$weight, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: waist**
```{r variable_waist, message = TRUE}
explainer(tb$waist)
```

```{r plot_waist}
plot(variable_plots$waist)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = waist)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$waist, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: cholesterol**
```{r variable_cholesterol, message = TRUE}
explainer(tb$cholesterol)
```

```{r plot_cholesterol}
plot(variable_plots$cholesterol)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = cholesterol)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$cholesterol, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: hdl**
```{r variable_hdl, message = TRUE}
explainer(tb$hdl)
```

```{r plot_hdl}
plot(variable_plots$hdl)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = hdl)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$hdl, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: ldl**
```{r variable_ldl, message = TRUE}
explainer(tb$ldl)
```

```{r plot_ldl}
plot(variable_plots$ldl)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = ldl)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$ldl, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: school**
```{r variable_school, message = TRUE}
explainer(tb$school)
```

```{r plot_school}
plot(variable_plots$school)
```

##### **Variable: family**
```{r variable_family, message = TRUE}
explainer(tb$family)
```

```{r plot_family}
plot(variable_plots$family)
```

##### **Variable: smoking**
```{r variable_smoking, message = TRUE}
explainer(tb$smoking)
```

```{r plot_smoking}
plot(variable_plots$smoking)
```

##### **Variable: myocard**
```{r variable_myocard, message = TRUE}
explainer(tb$myocard)
```

```{r plot_myocard}
plot(variable_plots$myocard)
```

##### **Variable: stroke**
```{r variable_stroke, message = TRUE}
explainer(tb$stroke)
```

```{r plot_stroke}
plot(variable_plots$stroke)
```

##### **Variable: diab_known**
```{r variable_diab_known, message = TRUE}
explainer(tb$diab_known)
```

```{r plot_diab_known}
plot(variable_plots$diab_known)
```

##### **Variable: diab_age**
```{r variable_diab_age, message = TRUE}
explainer(tb$diab_age)
```

```{r plot_diab_age}
plot(variable_plots$diab_age)
```

**Normality test**
```{r}
# QQ plot
ggplot(tb, aes(sample = diab_age)) + stat_qq(color='red', alpha = 0.6) + stat_qq_line() + theme_minimal()

# Normality assumption test
nt <- norm_test_fun(tb$diab_age, method = 'ks')
```

The `r nt$method` has a p-value of **`r round(nt$p.value, 4)`**.
Since `r ifelse(nt$p.value < 0.05, 'p-value is less than the significance level (0.05), we',
'p-value is not below the significance level (0.05), we do not have sufficient evidence to')`
reject the null hypothesis. Therefore, we can say that this variable **`r ifelse(nt$p.value < 0.05, 'does not follow', 'follows')`
the normal distribution**.

##### **Variable: contraception**
```{r variable_contraception, message = TRUE}
explainer(tb$contraception)
```

```{r plot_contraception}
plot(variable_plots$contraception)
```

##### **Variable: income**
```{r variable_income, message = TRUE}
explainer(tb$income)
```

```{r plot_income}
plot(variable_plots$income)
```

***



#### CORRELATION & ASSOCIATION 
In general there can be three types of association based on the data type of variables -
1. Between 2 continuous (numeric) variables
2. Between 2 categorical variables
3. Between 1 continuous and 1 categorical variables

In this section, each type will be analyzed separately. **association** function can be used to calculate these automatically.
The normality test of the variables will be done using the **'Kolmogorov-Smirnov Test'**.

```{r association}
corr_all <- association(tb, categorical = c('sex', 'obs_bp', 'obs_soma', 'obs_int', 'school', 'family', 'smoking', 'myocard', 'stroke', 'diab_known', 'contraception', 'income'), normality_test_method = 'ks')
```

##### Between 2 continuous (numeric) variables

For this case, the **association** function internally utilizes the **cor** function. Since there are some variables which may not follow the normality assumption, we will keep the method as 'auto'. This will assign the method automatically to all the variables pair based on normality tests.

```{r CC}
corrplot(corr_all$continuous_corr, p.mat = corr_all$continuous_pvalue, sig.level = .2)
```
The above correlations are either Pearson or Spearman, decided by the function based on the normality tests. A value close to 0 means no linear correlation between the variables. As values increases towards +1, the linear positive correlation increases (one variable increases as other increases). Similarly, as value decreases towards -1, the linear negative correlation increases (as one variable increases, other decreases). The insignificant correlation (based on p-value) are crossed out  at the desired level of significance (0.2).



##### Between 2 categorical variables

Association can be measured between categorical variables using the chi-square test of independence on two-way contingency tables. A small p-value indicates that association is present between the variables. Using the Chi-Square test statistic a measure of association can be calculated which is known as Cramer's V. This normalized measure varies from a minimum value of 0 when x and y are statistically independent to a maximum value of 1 when one variable is perfectly predictable from the other. Like correlation we can measure the magnitude of association using Cramer's V.

The Cramer's V can be calculated using the Chi-Square test statistic using:

$$V = \sqrt{\frac{\chi ^2}{n*min(k-1,r-1))}}$$

Where:
**$\chi ^{2}$** is derived from Pearson's chi-squared test
**$n$** is the grand total of observations
**$k$** is the number of columns
**$r$** is the number of rows.

```{r QQ}
corrplot(corr_all$categorical_cramers, p.mat = corr_all$categorical_pvalue, sig.level = .2)
```

One thing to note is that Cramer's V doesn't give direction of association using negative and positive values like Pearson / Spearman correlation. But this only give magnitude of association. The insignificant association (based on p-value) are crossed out  at the desired level of significance (0.2).

##### Between 1 continuous and 1 categorical variables

Association between 1 continuous and 1 categorical variable can be measured using the t-test (for normally distributed data) or Mann-Whitney (for data with distribution other than normal). A p-value smaller than the significant level indicates that both the variables are predictable by each other. Unlike previous two cases, these test doesn't give a magnitude of association between 0-1, but only a p-value showing the significance.

```{r CQ}
amat <- pmat <- corr_all$continuous_categorical
amat[amat > 0] <- 0
corrplot(amat, p.mat = pmat, sig.level = 0.2)
```

The insignificant association (based on p-value) are crossed out at the desired level of significance (0.2).

***

